Repeatability of Pre-Flashover Fire Patterns on Gypsum Wallboard

### 2017-02-27 Projects/Madrzykowski_Thesis/Report/Madrzykowski_Thesis.tex
minor changes

The Fire Dynamics Simulator (FDS) and the scientific visualization tool, Smokeview, are the most widely used and well documented CFD tools in the fire protection engineering community. Given the extent and high quality of the existing FDS documentation, there is no value to attempting to recreate as description of the model here other than to direct the reader to the volumes of information and data that exist for the FDS and Smookeview suite.

FDS~\cite{FDS_Users_Guide} is a CFD model developed and maintained by NIST that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires. A CFD model has a computational domain that is divided into  three-dimensional rectangular control volumes or computational cells. The size and arrangement of computational cells is sometimes referred to as the grid or the mesh.  The CFD model computes the density, velocity, temperature, pressure, and gas concentrations in each cell. Based on the laws of conservation of mass, momentum, and energy, the model tracks the generation and movement of fire gases. One of the most important aspects of FDS is that it is mathematically verified~\cite{FDS_Verification_Guide}and validated against fire test data~\cite{FDS_Validation_Guide}to ensure that it is accurate and provides the expected results, given appropriate input data. A complete description of the FDS model is provided in the FDS Technical Reference Guide~\cite{FDS_Tech_Guide}.

Smokeview is a software tool specifically designed to visualize simulation results from FDS~\cite{Smokeview_Users_Guide}. Smoke-view visualizes smoke and other attributes of the fire simulation by showing tracer particle flow and two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data, including temperature and flow vectors showing flow direction and magnitude. Smokeview has the ability to visualize fire and smoke by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS. Smokeview can also visualize data at particular snapshots in time using 2D or 3D contours of data that show temperature, flow direction, and flow magnitude.  Smokeview also has a feature that can represent a fire pattern on a surface based on user supplied threshold criteria.

@@

Two FDS models were developed: 1) to simulate the instrumented gypsum wallboard experiments with the natural gas burner and 2) to simulate the compartment fire pattern experiments with the natural gas burner.  One of the key determinations that is required when developing an FDS model is the mesh or grid cell size.  If a single plane of the data from the model were compared to a digital image we might refer to the grid cell size as the resolution of the image.  The smaller the grid cells the better the resolution. However increased resolution or the smaller grid cell size, the longer it will take the simulation to run.

>>>>

The Fire Dynamics Simulator (FDS) and the scientific visualization tool, Smokeview, are the most widely used and well documented CFD tools in the fire protection engineering community. In addition these tools have on going verification and validation programs.  Given the extent and high quality of the existing FDS documentation, there is no value to attempting to recreate a description of the model here other than to direct the reader to the volumes of information and data that exist for the FDS and Smookeview suite.

FDS~\cite{FDS_Users_Guide} is a CFD model initially developed and currently maintained by NIST that solves a form of the Navier-Stokes equations appropriate for low-speed, thermally driven flow with an emphasis on smoke and heat transport from fires. A CFD model has a computational domain that is divided into  three-dimensional rectangular control volumes or computational cells. The size and arrangement of computational cells is sometimes referred to as the grid or the mesh.  The CFD model computes the density, velocity, temperature, pressure, and gas concentrations in each cell. Based on the laws of conservation of mass, momentum, and energy, the model tracks the generation and movement of fire gases. One of the most important aspects of FDS is that it is mathematically verified~\cite{FDS_Verification_Guide}and validated against fire test data~\cite{FDS_Validation_Guide}to ensure that it is accurate and provides the expected results, given appropriate input data. A complete description of the FDS model is provided in the FDS Technical Reference Guide~\cite{FDS_Tech_Guide}.

Smokeview is a software tool specifically designed to visualize simulation results from FDS~\cite{Smokeview_Users_Guide}. Smokeview visualizes smoke and other attributes of the fire simulation by showing tracer particle flow and two dimensional (2D) or three dimensional (3D) shaded contours of gas flow data, including temperature and flow vectors showing flow direction and magnitude. Smokeview has the ability to visualize fire and smoke by displaying a series of partially transparent planes where the transparencies in each plane (at each grid node) are determined from soot densities computed by FDS. Smokeview can also visualize data at particular snapshots in time using 2D or 3D contours of data that show temperature, flow direction, and flow magnitude.  Smokeview also has a feature that can represent a fire or char pattern on a surface based on user supplied threshold criteria.

@@

Several FDS models were developed to correlate with the different types of experiments conducted over the course of this study: 1) simulations of the source fire characterization experiments documented in Chapter 2, 2) simulations of the instrumented gypsum wallboard experiments with the natural gas burner as documented in Chapter 4, and 3) simulations of the compartment fire pattern experiments with the natural gas burner in positions against a wall and in a corner as documented in Chapter 6.

The source fire simulation will be compared against the source fire characterization data to determine the best grid cell resolution and other inputs to use to provide the best fire simulation for use in the simulations of the wall and compartment experiments.

\subsection{Natural Gas Source Fires}

One of the key determinations that is required when developing an FDS model is the mesh or grid cell size.  If a single plane of the data from the model were compared to a digital image we might refer to the grid cell size as the resolution of the image.  The smaller the grid cells the better the resolution. However increased resolution or the smaller grid cell size, the longer it will take the simulation to run.
